Enforced Invariant Confluence

### 2016-11-22 doc/counters.tex
Started drawing rambling thoughts.
Added more to i-confluence alternatives.

A \emph{database state} is a total function $D: \var \to \ints$.  We say a
database state $D$ \emph{satisfies invariant} $I$ if $I$ evaluates to true
after all variables in $I$ have been replaced by their mapping in $D$.

@@

A transaction chain $C$ created from a set of transactions $T$ is a sequence of
transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
$C$ \emph{satisfies invariant} $I$ starting at $D$ if for every prefix $C'$ of
$C$, $D \circ C'$ satisfies invariant $I$.

@@

An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
is a sequence of transactions $c_1, \ldots, c_n$ where $c_i \in T$ for $1 \leq
i \leq n$. Each \imp{} transaction chain $B$ starting with database state $D$
has a corresponding fixed transaction chain $C$. Let $D_i = D \circ c_1 \circ
\cdots \circ c_i$. Then, the corresponding fixed transaction chain is
$\denote{c_1}(D_0), \ldots, \denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies
invariant} $I$ starting at $D$ if the corresponding fixed transaction chain $C$
satisfies invariant $I$ starting at $D$.

@@

>>>>

A \emph{database state} is a total function $D: \var \to \ints$.  We say a
database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
evaluates to true after all variables in $I$ have been replaced by their
mapping in $D$.

@@

A transaction chain $C$ created from a set of transactions $T$ is a sequence of
transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.

@@

An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
state $D$ has a corresponding fixed transaction chain $C$ denoted
$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
satisfies invariant $I$ starting at $D$.

@@

\subsection{\iconfluence{} Alternatives}
\newcommand{\ipreservation}{$\iinvariant$-preservation}
\newcommand{\ipreserving}{$\iinvariant$-preserving}
\newcommand{\isafety}{$\iinvariant$-safety}
\newcommand{\isafe}{$\iinvariant$-safe}

\begin{itemize}
  \item \todo{Clarify these rambling thoughts.}
  \item \todo{Add $k$-\isafety to the figure.}
  \item \todo{Find something between \isafety{} and \iconfluence.}
\end{itemize}

\begin{itemize}
  \item \textbf{1-\isafety.}
    $\forall D, D', c.\>
      I(D) \land
      I(D') \implies
      I(D \circ \denote{c}(D'))$

  \item \textbf{k-\isafety.}
    $\forall D, D', B, C, c.\>
        C = \denote{B}(D) \land
       |B| \leq k-1 \land
       I(C @ D) \land
       I(B @ D') \implies
       I(D \circ C \circ \denote{c}(D' \circ C))$

  \item \textbf{\isafety.}
    $\forall D, D', B, C, c.\>
        C = \denote{B}(D) \land
       I(C @ D) \land
       I(B @ D') \implies
       I(D \circ C \circ \denote{c}(D' \circ C))$

  \item \textbf{\ipreservation.}
    $\forall D, c.\>
       I(D) \implies
       I(D \circ \denote{c}(D)$

  \item \textbf{\iconfluence.}
    $\forall D, B_1, B_2, C_1, C_2. \>
       C_1 = \denote{B_1}(D) \land
       C_2 = \denote{B_2}(D) \land
       I(B_1 @ D) \land
       I(B_2 @ D) \implies
       I(D \circ C_1 \circ C_2)$

  \item \textbf{k-\iconfluence.}
    $\forall D, B_1, B_2, C_1, C_2. \>
       C_1 = \denote{B_1}(D) \land
       C_2 = \denote{B_2}(D) \land
       |B_1|, |B_2| \leq k \land
       I(B_1 @ D) \land
       I(B_2 @ D) \implies
       I(D \circ C_1 \circ C_2)$

  \item \textbf{1-\iconfluence.}
    $\forall D, c_1, c_2.\>
       I(D) \land
       I(D \circ \denote{c_1}(D)) \land
       I(D \circ \denote{c_2}(D)) \implies
       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
\end{itemize}

\begin{claim}
  $\text{\iconfluence} \implies \text{$k$-\iconfluence}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{1-\iconfluence} \centernot\implies \text{$k$-\iconfluence}$
\end{claim}
\begin{proof}
  See \clmref{wimp-1-iconfluence}.
\end{proof}

\begin{claim}
  $k\text{-\iconfluence} \implies \text{1-\iconfluence}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{$k$-\iconfluence} \centernot\implies \text{\iconfluence}$
\end{claim}
\begin{proof}
  See \clmref{wimp-1-iconfluence}.
\end{proof}

\begin{claim}
  $\text{\isafety} \centernot\implies \text{\ipreservation}$
\end{claim}
\begin{proof}
  \todo{Trivial.}
\end{proof}
\begin{claim}
  $\text{\ipreservation} \centernot\implies \text{\isafety}$
\end{claim}
\begin{proof}
  \todo{Not sure.}
\end{proof}

\tikzstyle{txn}=[ultra thick, ->]
\tikzstyle{dtxn}=[ultra thick, ->, dashed]
\tikzstyle{impl}=[-implies, double distance=3pt]
\newcommand{\inode}[3][]{
  \node[draw, shape=circle, minimum width=0.5cm, #1] (#2) at (#3) {};
  \node[fill, shape=circle] () at (#3) {};
}
\newcommand{\nnode}[3][]{
  \node[fill, shape=circle, #1] (#2) at (#3) {};
}

  \begin{tikzpicture}
    \node[] (isafety) at (0, 0) {\isafety};
    \node[] (ipreservation) at ($(isafety) + (30:3)$) {\ipreservation};
    \node[] (iconfluence) at ($(isafety) + (-30:3)$) {\iconfluence};
    \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
    \node[] (1-iconfluence) at ($(k-iconfluence) + (4, 0)$) {1-\iconfluence};

    \draw[impl] (isafety) -> (ipreservation);
    \draw[impl] (isafety) -> (iconfluence);
    \draw[impl] (iconfluence) -> (k-iconfluence);
    \draw[impl] (k-iconfluence) -> (1-iconfluence);
  \end{tikzpicture}

  \caption{}
  \label{}
\end{figure}

  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{d1}{0, 0}
      \inode[label=above:$D'$]{d2}{1, 0}
      \nnode{b1}{0, -1}
      \nnode{b2}{1, -1}
      \draw[txn] (d2) -- (b2) node[midway, right]{$t(D')$};
      \draw[dtxn] (d1) -- (b1) node[midway, left]{$t(D')$};
    \end{tikzpicture}
    \caption{1-\isafety}
    \label{fig:1-isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \nnode{b}{0, -1}
      \draw[txn] (a) -- (b) node[midway, left]{$t(D)$};
    \end{tikzpicture}
    \caption{1-\ipreservation}
    \label{fig:1-ipreservation}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{a}{0, 0}
      \inode[label=left:$D_{t}$]{l1}{$(a) + (240:1)$}
      \inode[label=right:$D_{u}$]{r1}{$(a) + (300:1)$}
      \nnode{b}{$(l1) + (300:1)$}
      \draw[txn] (a) -- (l1) node[midway, left] {$t(D)$};
      \draw[txn] (a) -- (r1) node[midway, right] {$u(D)$};
      \draw[dtxn] (l1) -- (b) node[midway, left] {$u(D)$};
      \draw[dtxn] (r1) -- (b) node[midway, right] {$t(D)$};
    \end{tikzpicture}
    \caption{1-\iconfluence}
    \label{fig:1-iconfluence}
  \end{subfigure}

  \vspace{1cm}

  \begin{subfigure}[b]{0.5\textwidth}
    \centering

    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D_0$]{l1}{0, 4}
      \inode[label=left:$D_1$]{l2}{0, 3}
      \inode[label=left:$D_{n-1}$]{l3}{0, 1}
      \nnode{l4}{0, 0}

      \inode[label=above:$D_0'$]{r1}{1, 4}
      \inode[label=right:$D_1'$]{r2}{1, 3}
      \inode[label=right:$D_{n-1}'$]{r3}{1, 1}
      \nnode{r4}{1, 0}

      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
      \draw[txn] (r2) -- (r3) node[midway, right]{
        $\begin{array}{l}
          \denote{c_2}(D_1'), \\
          \ldots, \\
          \denote{c_{n-1}}(D_{n-2}'), \\
        \end{array}$
      };
      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_n}(D_{n-1}')$};

      \draw[dtxn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
      \draw[dtxn] (l2) -- (l3) node[midway, left]{
        $\begin{array}{l}
          \denote{c_2}(D_1'), \\
          \ldots, \\
          \denote{c_{n-1}}(D_{n-2}'), \\
        \end{array}$
      };
      \draw[dtxn] (l3) -- (l4) node[midway, left]{$\denote{c_n}(D_{n-1}')$};
    \end{tikzpicture}

    \caption{\isafety}
    \label{fig:isafety}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering

    \begin{tikzpicture}[scale=2]
      \inode[label=above:{$D = D_{t_0} = D_{u_0}$}]{a}{0, 0}
      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1.5)$}
      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (0, -2)$}
      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1.5)$}
      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (0, -2)$}
      \nnode{b}{$(l2) + (300:1.5)$}

      \draw[txn] (a) -- (l1) node[midway, left] {$t_1(D_{t_0})$};
      \draw[txn] (a) -- (r1) node[midway, right] {$u_1(D_{u_0})$};
      \draw[txn] (l1) -- (l2) node[midway, left] {
        $\begin{array}{l}
          t_2(D_{t_1}), \\
          \ldots, \\
          t_n(D_{t_{n-1}})
        \end{array}$
      };
      \draw[txn] (r1) -- (r2) node[midway, right] {
        $\begin{array}{l}
          u_2(D_{u_1}), \\
          \ldots, \\
          u_m(D_{u_{m-1}})
        \end{array}$
      };
      \draw[dtxn] (l2) -- (b) node[midway, left] {
        $\begin{array}{l}
          u_i(D_{u_{i-1}}), \\
          1 \leq i \leq m
        \end{array}$
      };
      \draw[dtxn] (r2) -- (b) node[midway, right] {
        $\begin{array}{l}
          t_i(D_{t_{i-1}}), \\
          1 \leq i \leq n
        \end{array}$
      };
    \end{tikzpicture}

    \caption{\iconfluence}
    \label{fig:iconfluence}
  \end{subfigure}

  \caption{}
  \label{}
\end{figure}
